Seminarium TI 3700\\ Data Analysis on Twitter: An Overview (working title)

### 2014-03-11 authorinstr_2.tex
Tag based update 2

Our adaptation of the model to resource
recommendation simply exchanges the roles of resources and tags with respect to each
other. This model-based approach generates a set of factor matrices which resembles
a special case of the Tucker decomposition of a tensor. The tensor itself is not directly
induced by the data, but rather reflects a ranking over the resources for each user-tag
pair. Thus, it is important to note that this model is applicable only to the special case of
tag-based resource recommendation, and not for resource recommendation tasks with
other requirements.
The model is built by considering observations in the data of the form (u, r+ , r− , t),
where (u, r+ , t) is a triple which is found in the data (a positive example of resource
selection) and (u, r− , t) is a triple not found in the data (a negative example of re-
source selection). An iterative gradient-descent algorithm is employed to optimize a116
J. Gemmell et al.
ranking function that prefers positive examples in the data over negative ones. Each of
four related matrices is updated until convergence is found. The matrices represent the
U
T
factor-reduced components of the specialized tensor factorization M = Uk Rk
+Tk Rk
,
U
where Uk is the user factor matrix, Tk is the tag factor matrix, Rk is the resource factor
T
matrix with respect to users and Rk
is the resource factor matrix with respect to tags, k
is the selected number of factors, and M is the personalized resource-ranking tensor.
Generating a resource recommendation for a given user u and tag t is simply a
matter of referring to the appropriate user-tag column of the ranking tensor M . The
relevance score of a resource given a user-tag pair is calculated as: φ(u, {t}, r) =

\end{enumerate}

>>>>

This model is restricted in a way because its output is a set of matrices that illustrates a special case of the Tucker decomposition of a tensor. 
The tensor is not directly linked to the dataset, but rather implies a ranking. It is therefor only useful in the case of tag-based resource recommendation, 
and not for resource recommendations with other requirements.

To build the model, the data has to be observed in the form $(u, r+, r-, t)$ where $(u, r+, t)$ and $(u, r-, t)$ is a triple respectively
found and not found in the dataset. A ranking function is then used to prefer positive examples over negative ones. An iterative gradient-descent
algorithm is used to optimize this function. All four matrices will be updated until they converge. The matrices represent the factor-reduced components
of the specialised tensor factorization:

$M = U_k R_k^U + T_k R_k^T$

With 
\begin{enumerate}
 \item $U_k$ the user factorization
 \item $T_k$ is the tag factor matrix
 \item $R_k^U$ is the resource factor matrix with respect to the users
 \item $R_k^T$ is the resource factor matrix with respect to the tags
 \item $k$ selected number of factors
 \item $M$ the personalized resource-ranking tensor
\end{enumerate}

The relevance score of a resource can now be calculated by:

$\phi (u, \{t\}, r) = \sum_{i=1}^k U_k\{u\}\{i\}R_k^U\{r\}\{i\} + T_k\{t\}\{i\}R_k^T\{r\}\{i\}$	
\end{enumerate}
